---
title: "RAB234_Trpn_Ndufa2_RNAseq_report"
author: "Marion Hardy"
date: "2023-06-15"
output: 
  html_document:
    toc: true 
    theme: spacelab 
    highlight: monochrome
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, cache = TRUE, echo = FALSE, warning = F, cache.lazy = F)
knitr::opts_chunk$set(fig.width=10, fig.height=15) 

library(rjson)
library(tidyverse)
library(ggrepel)
library(clusterProfiler)
library(msigdbr)
library(readxl)

```


```{r Data loading, include=FALSE}

res_tbl <- read_csv("./data_output/Trpn_Nduf_vs_Scr/res_tbl.csv")
diff <- read_csv("./data_output/Trpn_Nduf_vs_Scr/res_tbl_signif.csv")
res_tbl1 <-  res_tbl %>% filter(!is.na(ENTREZID), padj<=0.05)

```

# Introduction

Data from  RAB : CD8+ T cells from mice

- Cell types: not nucleofected, scramble, Ndufa2 KO, Mrpl28 KO

-   Conditions: treated w/ or w/o Tryptophan (4 uM vs 70 uM)

-   Biological replicates (3)

# Volcano plot 

```{r, fig.height = 6, fig.width = 7}

res_tbl %>%
  filter(!is.na(padj)) %>%
  ggplot(aes(x = log2FoldChange, y = -log10(padj),
             color = padj < 0.05 & abs(log2FoldChange) > 1, 
             label = ifelse(padj<0.05&log2FoldChange>=1|
                              padj<0.05&log2FoldChange<=-1,as.character(gene),''))) +
  scale_colour_manual(values = c("gray", "firebrick3")) +
  labs(col="Significantly expressed")+
  geom_point(size = 0.5) +
  geom_hline(yintercept = -log10(0.05)) +
  geom_vline(xintercept = 1) +
  geom_vline(xintercept = -1)+
  geom_text_repel(max.overlaps = 15,
                  box.padding = 0.25,
                  segment.color = 'grey50',
                  fontface = "italic")+
  labs(title = "Ndufa2 KO vs scramble cells in Trpn medium")+
  theme_bw()

```

Getting some stats yay

```{r, echo=TRUE}

table(res_tbl$padj<=0.05) # significant genes

table(res_tbl$padj<=0.05&res_tbl$log2FoldChange<=-1) # signif and downreg

table(res_tbl$padj<=0.05&res_tbl$log2FoldChange>=1) # signif and upreg

```

# GSEA

```{r}

ordered_genes_fc <- res_tbl1$log2FoldChange
names(ordered_genes_fc) <- res_tbl1$gene
ordered_genes_fc <- sort(ordered_genes_fc, decreasing = T)

```

```{r}

mm_GO_sets <- msigdbr(
  species = "Mus musculus", 
  category = "C5",
  subcategory = "BP") # for gene ontology collection

mm_reactome_sets <- msigdbr(
  species = "Mus musculus", 
  category = "C2",
  subcategory = "CP:REACTOME") # for reactome collection

mm_wiki_sets <- msigdbr(
  species = "Mus musculus", 
  category = "C2",
  subcategory = "CP:WIKIPATHWAYS") # for Wikipathways collection

mm_pert_sets <- msigdbr(
  species = "Mus musculus", 
  category = "C2",
  subcategory = "CGP")

mm_hallmark_sets <- msigdbr(
  species = "Mus musculus", 
  category = "H") # for hallmark collection

mm_GO_gl = 
  mm_GO_sets %>% 
  filter(grepl("glycoly", gs_name, ignore.case = T)) %>% 
  rbind(mm_hallmark_sets %>% filter(grepl("glycoly", gs_name, ignore.case = T))) %>% 
  rbind(mm_reactome_sets %>% filter(grepl("glycoly", gs_name, ignore.case = T)))

mm_GO_ox  = 
  mm_GO_sets %>% 
  filter(grepl("oxidative", gs_name, ignore.case = T)|
         grepl("respiration", gs_name, ignore.case = T)) %>% 
  rbind(mm_hallmark_sets %>% filter(grepl("oxidative", gs_name, ignore.case = T),
         grepl("respiration", gs_name, ignore.case = T)) %>% 
  rbind(mm_reactome_sets %>% filter(grepl("oxidative", gs_name, ignore.case = T))))

```


## Gene ontology

```{r, fig.height=5, fig.width=7}

gsea_results_go <- GSEA(
  geneList = ordered_genes_fc, 
  pvalueCutoff = 0.5, 
  eps = 0, 
  pAdjustMethod = "BH", 
  nPermSimple = 10000,
  TERM2GENE = dplyr::select(
    mm_GO_sets,
    gs_name,
    gene_symbol
  )
)

# dotplot(gsea_results_go, x = "NES", showCategory = 30, font.size = 8)+ 
#   ggtitle("GSEA LFC Gene ontology")
# 
# gsea_results_go %>% 
#   as_tibble() %>% 
#   select(ID, setSize,	enrichmentScore,	NES,	pvalue,	p.adjust,	qvalue) %>%
#   knitr::kable()

```

No results

## Reactome

```{r, fig.height=5, fig.width=7}

gsea_results_reac <- GSEA(
  geneList = ordered_genes_fc, 
  pvalueCutoff = 0.5, 
  eps = 0, 
  seed = TRUE, 
  pAdjustMethod = "BH", 
  nPermSimple = 10000, 
  TERM2GENE = dplyr::select(
    mm_reactome_sets,
    gs_name,
    gene_symbol
  )
)

# dotplot(gsea_results_reac, x = "NES", showCategory = 30, font.size = 8)+ 
#   ggtitle("GSEA LFC Reactome")
# 
# gsea_results_reac %>%
#   as_tibble() %>%
#   select(ID, setSize,	enrichmentScore,	NES,	pvalue,	p.adjust,	qvalue) %>%
#   knitr::kable()
```

No results

## Wikipathways

```{r, fig.height=5, fig.width=7}

gsea_results_wiki <- GSEA(
  geneList = ordered_genes_fc, 
  pvalueCutoff = 0.5, 
  seed = TRUE, 
  pAdjustMethod = "BH", 
  TERM2GENE = dplyr::select(
    mm_wiki_sets,
    gs_name,
    gene_symbol
  )
)

# dotplot(gsea_results_wiki, x = "NES", showCategory = 30, font.size = 8)+
#   ggtitle("GSEA LFC Wikipathways")
# 
# gsea_results_wiki %>%
#   as_tibble() %>%
#   select(ID, setSize,	enrichmentScore,	NES,	pvalue,	p.adjust,	qvalue) %>%
#   knitr::kable()

```

No results

## Perturbations

```{r, fig.height=5, fig.width=7}

gsea_results_pert <- GSEA(
  geneList = ordered_genes_fc, 
  pvalueCutoff = 0.5, 
  seed = TRUE, 
  pAdjustMethod = "BH", 
  TERM2GENE = dplyr::select(
    mm_pert_sets,
    gs_name,
    gene_symbol
  )
)

# dotplot(gsea_results_pert, x = "NES", showCategory = 30, font.size = 8)+
#   ggtitle("GSEA LFC genetic perturbations")
# 
# gsea_results_pert %>%
#   as_tibble() %>%
#   select(ID, setSize,	enrichmentScore,	NES,	pvalue,	p.adjust,	qvalue) %>%
#   head(15) %>%
#   knitr::kable()

```

No results

## Hallmarks

```{r, fig.height=5, fig.width=7}

gsea_results_hall <- GSEA(
  geneList = ordered_genes_fc,
  pvalueCutoff = 0.5, 
  seed = TRUE, 
  pAdjustMethod = "BH", 
  TERM2GENE = dplyr::select(
    mm_hallmark_sets,
    gs_name,
    gene_symbol
  )
)

# dotplot(gsea_results_hall, x = "NES", showCategory = 30, font.size = 8)+ 
#   ggtitle("GSEA LFC hallmark")
# 
# gsea_results_hall %>% 
#   as_tibble() %>% 
#   select(ID, setSize,	enrichmentScore,	NES,	pvalue,	p.adjust,	qvalue) %>%
#   knitr::kable()

```

No results

# ORA
## Gene ontology

```{r, fig.height = 5, fig.width = 7}

bckgrd = res_tbl1 %>% filter(abs(log2FoldChange)>=1)
  
up = res_tbl1 %>% filter(abs(log2FoldChange) >= 1)
up = up$gene

ora_results <- enricher(
  gene = up, 
  pvalueCutoff = 0.05, 
  pAdjustMethod = "BH", 
  minGSSize = 10,
  TERM2GENE = dplyr::select(
    mm_GO_sets,
    gs_name,
    gene_symbol
  )
)

dotplot(ora_results, showCategory = 30, font.size = 8)+ 
  ggtitle("Over-representation analysis: Gene Ontology")

ora_results %>%
  as_tibble() %>%
  select(ID, BgRatio, p.adjust, qvalue, geneID) %>%
  knitr::kable()

```


## Reactome

```{r, fig.height = 5, fig.width = 7}

ora_results <- enricher(
  gene = up, 
  pvalueCutoff = 0.05, 
  pAdjustMethod = "BH", 
  minGSSize = 10,
  TERM2GENE = dplyr::select(
    mm_reactome_sets,
    gs_name,
    gene_symbol
  )
)

# dotplot(ora_results, showCategory = 30, font.size = 8)+ 
#   ggtitle("Over-representation analysis: Reactome")
# 
# ora_results %>%
#   as_tibble() %>%
#   select(ID, BgRatio, p.adjust, qvalue, geneID) %>%
#   knitr::kable()

```

No results

## Hallmark

```{r, fig.height = 6, fig.width = 7}

ora_results <- enricher(
  gene = up, 
  pvalueCutoff = 0.05, 
  pAdjustMethod = "BH", 
  minGSSize = 10,
  TERM2GENE = dplyr::select(
    mm_hallmark_sets,
    gs_name,
    gene_symbol
  )
)

dotplot(ora_results, showCategory = 30, font.size = 8)+ 
  ggtitle("Over-representation analysis: Hallmark")

ora_results %>%
  as_tibble() %>%
  select(ID, BgRatio, p.adjust, qvalue, geneID) %>%
  knitr::kable()

```

## Wikipathways

```{r, fig.height = 6, fig.width = 7}

ora_results <- enricher(
  gene = up, 
  pvalueCutoff = 0.05, 
  pAdjustMethod = "BH", 
  minGSSize = 10,
  TERM2GENE = dplyr::select(
    mm_wiki_sets,
    gs_name,
    gene_symbol
  )
)

# dotplot(ora_results, showCategory = 30, font.size = 8)+ 
#   ggtitle("Over-representation analysis: Wikipathways")
# 
# ora_results %>%
#   as_tibble() %>%
#   select(ID, BgRatio, p.adjust, qvalue, geneID) %>%
#   knitr::kable()

```

No results

## Perturbations

```{r, fig.height = 10, fig.width = 7}

ora_results <- enricher(
  gene = up, 
  pvalueCutoff = 0.05, 
  pAdjustMethod = "BH", 
  minGSSize = 10,
  TERM2GENE = dplyr::select(
    mm_pert_sets,
    gs_name,
    gene_symbol
  )
)

dotplot(ora_results, showCategory = 30, font.size = 8)+ 
  ggtitle("Over-representation analysis: Perturbations")

ora_results %>%
  as_tibble() %>%
  select(ID, BgRatio, p.adjust, qvalue, geneID) %>%
  head(30) %>% 
  knitr::kable()

```


# Pattern search in data
## Slcxxx transporters

```{r, fig.height = 6, fig.width = 7}

slc_pattern = res_tbl1[grep("slc",res_tbl1$gene, ignore.case = T),]

res_tbl %>%
  filter(!is.na(padj)) %>%
  ggplot(aes(x = log2FoldChange, y = -log10(padj),
             color = padj < 0.05 & abs(log2FoldChange) > 1, 
             label = ifelse(padj<0.05&log2FoldChange>=1|
                              padj<0.05&log2FoldChange<=-1,as.character(gene),''))) +
  scale_colour_manual(values = c("gray", "lightcoral")) +
  geom_point(size = 0.5) +
  geom_hline(yintercept = -log10(0.05)) +
  geom_vline(xintercept = 1) +
  geom_vline(xintercept = -1)+
  theme_bw()+
  geom_point(data=slc_pattern, 
             aes(x=log2FoldChange,y=-log10(padj)), 
             color='blue',
             size=0.5)+
  geom_text_repel(data = slc_pattern, size = 3, segment.color = "blue",
                  max.overlaps = 30, min.segment.length = 0, color = "blue",
                  box.padding = 0.5)+
  theme(legend.title= element_blank())


slc_pattern%>% 
  select(gene, log2FoldChange, padj, description) %>% 
  arrange(gene) %>% 
  knitr::kable()
```

## Mitochondria mention in gene description

```{r, fig.height = 6, fig.width = 7}

mitoc_pattern = res_tbl1[grep("mitoc",res_tbl1$description, ignore.case = T),]

res_tbl %>%
  filter(!is.na(padj)) %>%
  ggplot(aes(x = log2FoldChange, y = -log10(padj),
             color = padj < 0.05 & abs(log2FoldChange) > 1, 
             label = ifelse(padj<0.05&log2FoldChange>=1|
                              padj<0.05&log2FoldChange<=-1,as.character(gene),''))) +
  scale_colour_manual(values = c("gray", "lightcoral")) +
  geom_point(size = 0.5) +
  geom_hline(yintercept = -log10(0.05)) +
  geom_vline(xintercept = 1) +
  geom_vline(xintercept = -1)+
  theme_bw()+
  geom_point(data=mitoc_pattern, 
             aes(x=log2FoldChange,y=-log10(padj)), 
             color='blue',
             size=0.5)+
  geom_text_repel(data = mitoc_pattern, size = 3, segment.color = "blue",
                  max.overlaps = 30, min.segment.length = 0, color = "blue",
                  box.padding = 0.5)+
  theme(legend.title= element_blank())


mitoc_pattern%>% 
  select(gene, log2FoldChange, padj, description) %>%
  arrange(gene) %>% 
  knitr::kable()

```
